---
title: "Clustering and classification"
author: "Sheila"
date: "18/11/2020"
output: html_document
---
**Load data **

Boston dataset contains data about housing in the area of Boston Mass as collected by the U.S Census Service.
```{r}
# Access MASS library 
library(MASS)

# Load Boston data
data("Boston")
```

```{r}
# explore Boston dataset
str(Boston)
dim(Boston)
```
The Boston dataset has has 506 rows and 14 columns. 

The summary of the dataset
```{r}
summary(Boston)

# plot matrix of the variables
pairs(Boston)

```


We can assess the correlations between variables with correlation matrix plot:

```{r}
library(corrplot)
library(tidyverse)

# First calculate the correlation matrix 
corr_matrix<-cor(Boston)%>% round(2)

# print corr_matrix
corr_matrix

# visualize the correlation matrix
corrplot(corr_matrix, method="circle", type = "upper", tl.cex = 1.2) %>% round(2)
```

Correlation Matrices
Correlation of target variable with predictor variables:

1. crim are highly correlated to indu, nox, age, rad, and tax and Istat and negatively to zn, rm,dis and medv
2. rm and lstat are highly correlated with the target variable medv
3. black, dis, rm, chas, zn are positively correlated with medv
4. crim, indus, nox, age, rad, tax, ptratio, lstat are negatively correlated with medv