# **Regession and Model Validation**

Sheila Wachiye

November 5 2020


# *Introduction*
What is regression?

Regression is a statistical method used to assess the relationship between a dependent variable and one and more independent variables.

What is model validation?


# *Steps*
1.Load "students2014" data to be analysized 

Do this by assigning it a name and in this case we use "lrn14".
We then use *read.table* function, to read the file in a table format and creates a data frame.

The separator is a comma "," and the file includes a header
```{r}
lrn14 <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/learning2014.txt", sep=",", header=TRUE)
```

2. To have an idea of how the data looks like, we can 

 i. Look at the data structure 
 This shows 
```{r}
str(lrn14)
```
 ii. or the data dimensions 
```{r}
dim(lrn14)
```
3. We can also Visualization the data by creating graphics using *ggplots function* in R

 Ggplot2 is an implementation of Leland Wilkinson's Grammar of Graphics which is popula for data  visualization.
Install ggplots using *install.packages(ggplots2)*

# Access the gglot2 using library
```{r}
library(ggplot2)
```

# To plot points against attitude,initialize plot with data and aesthetic mapping. 
Include *col(Gender)* to highlight the gender distribution based on the attitude.
```{r}
p1 <- ggplot(lrn14, aes(x = attitude, y = points, col= gender))
p2 <- p1 + geom_point()
p2
```









