
#**Principal Component Analysis **
```{r}
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep=",", header=TRUE)

# Accesss necessary library
library(dplyr)
library(ggplot2)
library(GGally)

```

Data structure

```{r}
# Structure 
str(human)

# Dimension
dim(human)
head(human)

```

Summarise Data

```{r}
summary(human)
```


```{r}
# glimpse at the human data
glimpse(human)

```
  
Explore relationships

```{r}
# Access GGally
library(GGally)

# To visualise the correlation between the variables in our data
ggpairs(human)

```


The diagonal plots give the density plot. The scale of the y axis is printed on the left side of the matrix and x axis is at the bottom of the matrix. 

The upper part of the density plot gives the pearson’s correlation coefficient of determination (R2) showing positive and negative (-) correlations between the variable in the human development Index. 

Variable that are significantly positive or significantly negative are give by the symbol (***)

The plot below the density plot shows the scatter plot between two variables


We can also look at the correlations using the cor from the *corrplot function*

```{r}
# Access corrplot
library(corrplot)

# compute the correlation matrix and visualize it with corrplot
cor(human) %>% corrplot

```



**Principal Component Analysis (PCA) **

Principal Component Analysis a method used to reduce the dimensionality of large data sets into a smaller one that still contains most of the information in the large set.

The number of PCs is equal to the number of original variables and for our case we have eight (8).

The aim is to keep only the PCs that explain the most variance of our data to make the interpretation easier.


**PCA on Unstandardise data**
```{r}
# perform principal component analysis (with the SVD method)
pca_human1 <- prcomp(human)
summary(pca_human1)

# draw a biplot of the principal component representation and the original variables
biplot(pca_human1, choices = 1:2, cex = c(0.8, 1),col = c("grey30", "red"))
```

```{r}

# Access the factoextra 
library(factoextra)

#Visualize eigenvalues 
fviz_eig(pca_human1)


# Another way to draw a biplot of the principal component
# Gives percentage variance of PC1 on the x axis and PC2 on the y axis
fviz_pca_biplot(pca_human1, repel = TRUE, col.var = "#2E8FDF", # Variables color
                col.ind = "#696959"  # Individuals color
                )

eig.val <- get_eigenvalue(pca_human1)
eig.val
```

Standardization is a advisable for data transformation especially because the variables in original dataset have different scale or different units. 

The standardized variables will therefore be unitless and have a similar variance.

This is necessary to eliminate biases in the original variables. 



**PCA on Standardise data**

```{r}
# standardize the variables
human_std <- scale(human)

# perform principal component analysis (with the SVD method)
pca_human2 <- prcomp(human_std)

#The summary shows outputs the proportion of variance explained by the components.
summary(pca_human2)

# Access the factoextra 
library(factoextra)

#Get eigenvalues to determine the number of principal components to be considered.
eig.val <- get_eigenvalue(pca_human2)
eig.val

#Visualize eigenvalues
#Scree plot,

fviz_eig(pca_human2)

```
The PCA results for standardised data are are different from unstandardized data.

PCA on standardised data are easier to interprete and see the cooreltion as compared to unstandardised data.

Based on these results

The PC1 account for 53.6% of the total variance and the PC2 16.2%. Thus PC1 and PC2 contributes about 69% of the total variation in the dataset and have eigenvalues > 1 
This mean, PC1 and PC2 provides a good approximation of the variation present in the original human dataset that has 8 variable.


```{r}

# draw a biplot of the principal component representation and the original variables
biplot(pca_human2, choices = 1:2, cex = c(0.8, 1), col = c("grey30", "red"))



# Anothe way to draw a biplot of the principal component
# Gives percentage variance of PC1 on the x axis and PC2 on the y axis
fviz_pca_biplot(pca_human2, repel = TRUE, col.var = "#2E5FDF", # Variables color
                col.ind = "#696959"  # Individuals color
                )
```

From the biplot, 
we can see the variables D and E are highly associated and forms cluster (gene expression response in D and E conditions are highly similar). Similarly, A and B are highly associated and forms another cluster (gene expression response in A and B conditions are highly similar but different from other clusters). If the variables are highly associated, the angle between the variable vectors should be as small as possible in the biplot.

