# **Regession and Model Validation**

Sheila Wachiye

November 5 2020


# *Introduction*
What is regression?

Regression is a statistical method used to assess the relationship between a dependent variable and one and more independent variables.

What is model validation?


# *Steps*
1.Load "students2014" data to be analysized 

Do this by assigning it a name and in this case we use "lrn14".
We then use *read.table* function, to read the file in a table format and creates a data frame.

The separator is a comma "," and the file includes a header
```{r}
lrn14 <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/learning2014.txt", sep=",", header=TRUE)
```

2. To have an idea of how the data looks like. 

We can look at the data structure 
```{r}
str(lrn14)
```
 We can also look at the data dimensions 
```{r}
dim(lrn14)
```

3. We can also Visualization the data by creating graphics using *ggplots function* in R

 Ggplot2 is an implementation of Leland Wilkinson's Grammar of Graphics which is popula for data  visualization.
Install ggplots using *install.packages(ggplots2)*

Access the *gglot2* using library
```{r}
library(ggplot2)
```

To plot points against attitude,initialize plot with data and aesthetic mapping. 

To highling the distribution of the gender based on the attitide, use *col(gender)*.
```{r}
p1 <- ggplot(lrn14, aes(x = attitude, y = points, col= gender))

# define the visualization type (points)
p2 <- p1 + geom_point()
# draw the plot
p2
# add the regression line
p3 <- p2 + geom_smooth(method = "lm")
# draw the plot with 
p3

# add title and draw the plot
p4 <- p3 + ggtitle("Student's attitude against exam points")
p4
```

# We can loodraw a scatter plot matrix of the variables in learning2014.
```{r}
# [-1] excludes the first column (gender)
 pairs(lrn14[-1])

# access the GGally and ggplot2 libraries
library(GGally)

# create a more advanced plot matrix with ggpairs()
p <- ggpairs(lrn14, mapping = aes(), lower = list(combo = wrap("facethist", bins = 20)))

# draw the plot
p
```

```{r}
# learning2014 is available

# a scatter plot of points versus attitude
library(ggplot2)
qplot(attitude, points, data = lrn14) + geom_smooth(method = "lm")

# fit a linear model
my_model <- lm(points ~ attitude, data = lrn14)

# print out a summary of the model

summary(my_model)
```
```{r}
# learning2014 is available

# a scatter plot of points versus attitude
library(ggplot2)
qplot(attitude, points, data = lrn14) + geom_smooth(method = "lm")

# fit a linear model
my_model1 <- lm(points ~ stra, data = lrn14)

# print out a summary of the model

summary(my_model1)
```
```{r}
# learning2014 is available

# a scatter plot of points versus attitude
library(ggplot2)
qplot(attitude, points, data = lrn14) + geom_smooth(method = "lm")

# fit a linear model
my_model <- lm(points ~ surf, data = lrn14)

# print out a summary of the model

summary(my_model)
```
```{r}
# learning2014 is available

# a scatter plot of points versus attitude
library(ggplot2)
qplot(attitude, points, data = lrn14) + geom_smooth(method = "lm")

# fit a linear model
my_model2 <- lm(points ~ attitude+ stra+ surf, data = lrn14)

# print out a summary of the model

summary(my_model2)
```